Generation of TypeScript Declaration Files from JavaScript Code

### 2020-08-27 programming-2021/main.tex
section 0-4

\title{Generation of TypeScript Declaration Files from JavaScript Code}

% \titlerunning{Dummy short title}%optional, please use if title is longer than one line

@@

\author{Peter Thiemann}
\authorinfo{\email{thiemann@informatik.uni-freiburg.de}}
\affiliation{Universität Freiburg, Germany}

@@

%TODO mandatory: add short abstract of the document
\begin{abstract}
Developers are starting to write large and complex applications in
TypeScript, a typed dialect of JavaScript. TypeScript applications
integrate JavaScript libraries via typed descriptions of their APIs
called declaration files. DefinitelyTyped is the standard public
repository for these files.
The repository is maintained manually by volunteers, which
is error prone and time consuming. Discrepancies between a
declaration file and the JavaScript implementation lead to
incorrect feedback from the TypeScript IDE and thus to incorrect uses
of the underlying JavaScript library.

This work presents \texttt{dts-generate}, a tool that generates
TypeScript declaration files for JavaScript libraries uploaded to the \NPM{}
registry. It extracts code examples from the documentation written by
the developer, executes the library driven by the examples, gathers
run-time information, and generates a declaration file based on this
information. To evaluate the tool, 244 declaration files were generated and 82 were
compared with the corresponding declaration file provided on DefinitelyTyped. 52 files presented positive results, having no differences at all or differences that can be solved by modifying the code examples accordingly.

\end{abstract}

@@

\begin{lstlisting}[
    caption={\textbf{Unintuitive JavaScript behavior} Falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float=bp,
    captionpos=b
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
[1] + 1; //'11'
[2] == "2"; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3'
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
typeof null; // object
null instanceof Object; // false
"01" < "00100"; // false
\end{lstlisting}
% pjt: why unexpected?

@@

Unfortunately, creation and maintenance for most declaration
files in DefinitelyTyped is conducted manually,
which is error prone and time consuming. As TypeScript takes a
declaration file at face value, discrepancies are not detected at
compile time and the resulting inaccuracies in autocompletion, say,
are confusing for the programmer.
As TypeScript does not perform any run-time 
checking of types, either,  discrepancies  between the declaration file
and its corresponding JavaScript library can lead to unexpected 
behavior and crashes. Such an experience can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

@@

We aim to improve on this situation by providing a tool that generates
sound TypeScript declaration files automatically. To do so, we rely on the examples provided
as part of the documentation of a library. If these
examples are sufficiently rich, then our toolchain can generate
high-quality declaration files from them. This approach takes
advantage of 
best practices in the dynamic languages community which favors
examples and tests over writing out type signatures.

As an example, consider the NPM module
\texttt{abs} that
``computes the absolute path of an input''. Its documentation consists
of the following three examples\footnote{\url{https://www.npmjs.com/package/abs}}.
\begin{lstlisting}[language=JavaScript,numbers=none]
const abs = require("abs");

@@

\item A comparator for TypeScript declaration files (Section~\ref{sec:dts-compare}). This tool is
  necessary for evaluating our framework and also useful to detect
  incompatibilities when evolving JavaScript modules.
\item An evaluation of our framework (Sections~\ref{sec:dts-generate-evaluation}
  and~\ref{sec:results}). We examined all 6029 entries in 
  the DefinitelyTyped repository and found 244 sufficiently
  well-documented \NPM{} packages, on which we ran \texttt{dts-generate}
  and compared the outcome with the respective declaration file from the
  DefinitelyTyped repository. 
\end{itemize}

\section{Motivating Example}
\label{sec:motivating-example}
The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a
glob expression for matching filenames in the shell into a regular
expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It 
has about 9,900,000 weekly downloads and 188 \NPM{} packages depend on it. If
a developer creates or extends TypeScript code that depends on the
\texttt{glob-to-regexp} library, the TypeScript
compiler and IDE requires a declaration file for that library to
perform static checking and code completion, respectively. With
\texttt{dts-generate} we automatically generate a TypeScript
declaration file for \texttt{glob-to-regexp}. The tool downloads the \NPM{} 
package, runs the examples extracted from its documentation, gathers
run-time information, and generates a TypeScript declaration
file. The result is 
shown in \figref{fig:motivating-example-glob-to-regexp-vscode} and it is
ready for use in a TypeScript project. For example,  Visual Studio's
code completion runs properly with the generated file
(\figref{fig:motivating-example-glob-to-regexp-vscode}). If the
\texttt{glob-to-regexp} package gets modified in the future, a new declaration
file can be generated automatically using
\texttt{dts-generate}. Our comparator tool can compare the new file
for incompatibilities with the previous declaration file.

@@

The declaration file shown in \figref{fig:motivating-example-glob-to-regexp-vscode}
describes a package with a single exported function. The first parameter is of type
\texttt{string} and the second one is an optional interface. TypeScript namespaces
organize types declared in a declaration file and avoid name clashes with other types
\cite{typescript-namespaces}. For example, the interface \texttt{I\_\_opts} declared in
\figref{fig:motivating-example-glob-to-regexp-vscode} is used as
\lstinline[language=TypeScript]{globToRegexp.I__opts} because it is declared within a
namespace.

This file is an instance of one of the standard templates for writing
declaration
files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}:
\textbf{module}, \textbf{module-class}, and
\textbf{module-function}. 
Each template corresponds to a different way of describing the exports
of a JavaScript library. The choice of the template depends on the
organization of the underlying JavaScript library:
\begin{description}
\item[module] several exported functions,
\item[module-class] a class-like structure,
\item[module-function] exactly one exported function.
\end{description}

\begin{lstlisting}[
    caption={Example for template \textbf{module-function}},
    language=bash,
	label=code:dts-generate-example,
    float,captionpos=b
]
$ ./dts-generate abs
$ cat output/abs/index.d.ts 
export = Abs;

@@

greet({
  greeting: "hello world",
  color: "#00ff00"
});
    \end{lstlisting}
    \caption{Example for \texttt{module-function} template}
    \label{fig:example-module-function}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=TypeScript,numbers=none]
export = GreetSettingsModule;

@@

}
    \end{lstlisting}
    \caption{Declaration file for \texttt{module-function} template}
    \label{fig:template-module-function}
  \end{subfigure}

@@

var goodbye = myLib.makeGoodBye();
console.log("The computed goodbye is: " + goodbye);    
      \end{lstlisting}
    \caption{Example for \texttt{module} template}
    \label{fig:example-module}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript,numbers=none]
export function makeGreeting(str: string): string;
export function makeGoodBye(): string;        
      \end{lstlisting}
      \caption{Declaration file for \texttt{module} template}
      \label{fig:template-module}
    \end{subfigure}

@@

var myGreeter = new Greeter("hello, world");
myGreeter.greeting = "howdy";
myGreeter.showGreeting();
      \end{lstlisting}
      \caption{Example for \texttt{module-class} template}
      \label{fig:example-class}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript,numbers=none]
export = Greeter;

@@

declare namespace Greeter {
}
      \end{lstlisting}
      \caption{Declaration file for \texttt{module-class} template}
      \label{fig:template-class}
    \end{subfigure}

@@

Like \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
file, we determine the kind of declaration file by examining the provided example. If the
imported entity is used solely as a function as in Figure~\ref{fig:example-module-function}, the
generated declaration file follows the module-function template as shown in Figure~\ref{fig:template-module-function}.
If the example only accesses properties of the imported entity as in
Figure~\ref{fig:example-module}, we generate according to the module template (Figure~\ref{fig:template-module}).
If the imported entity is used with \lstinline/new/ to create new instances as in
Figure~\ref{fig:example-class}, we generate a module-class template (Figure~\ref{fig:template-class}).

The libraries \texttt{glob-to-regexp} as well as \texttt{abs} are both instances of the
\textbf{module-function} template.
Most libraries fall in one of these three categories. 
While it is possible to create a library of undetermined category, our selection is driven
by the examples in the documentation and 
hence reflects the developers intent.
% Run-time analysis proves useful in this case,
% since it is not sufficient to statically analyze the exported module.

@@

The examples and the code base of the library are instrumented with
Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
information and type information at run time. Jalangi is a configurable
framework for dynamic analysis of JavaScript. It provides several
analysis modules that we extended as needed to retrieve the required
run-time information, which is then saved in a JSON file.

A second independent block uses the run-time information to generate a
TypeScript declaration file. It infers the overall structure of the JavaScript
library, its interfaces, and the types from the run-time information. 
The resulting declaration file is ready for use in the development
process. Its contents mimic the usage of the library in the example
code fragments and matches the structure of the
JavaScript library under analysis, so that the JavaScript code
generated after compiling the TypeScript code runs without
interface-related errors.

The command-line interface is inspired by the \texttt{dts-gen} tool
\cite{dts-gen}. \coderef{code:dts-generate-example} shows
that invoking the package is very simple. Its only required
argument is the name of the module published to the \NPM{} registry.

@@

To extract run-time information from a JavaScript library, it is
necessary, by definition, to actually execute the code, because the
analysis modules provided by Jalangi to gather information are only
triggered if the instrumented code gets executed.

@@

Option~\ref{item:1} does not solve our problem, it just delegates it
to the importing library, which also needs code to drive it. Moreover,
it is costly to download and instrument another package.

We considered option~\ref{item:2} under the assumption that most
libraries would come with test cases. However, there is no standard
for testing JavaScript code so that test cases were difficult to reap
from the \NPM{} packages: they use different directory structures,
employ differing (or no) testing tools, or do not have tests
at all. Moreover, developers try to cover edge cases or to trigger errors by using illegal
input values in their tests. Such tests are not really useful for describing a library
from the consumer's point of view. For example, a developer might write a test invoking a
function with \texttt{null} just to validate that it throws an error in such scenario.

In the end, option~\ref{item:3} was the most viable even though there
is no standard for documentation, either. However, almost all repositories
contain README files where the library authors briefly describe in prose what
the code does, which problem it solves, how to install the
application, how to build the code, etc. It is very common that
developers provide code examples in the README files to show how the
library works and how to use it. These code fragments represent common usage cases rather
than stress-testing the implementation (the problem of
option~\ref{item:2}) because they are meant to be instructive to users of the package.

@@

\subsection{Run-time Information Gathering}
\label{sec:run-time-information}
The Runtime Information block described in
\figref{fig:tsd_generation_method_block_diagram} gathers
type-related information as well as usage-related information for each function parameter.
For example,

@@

Instead of analyzing the shape of the exported module, we inspect how the module is used
to choose between the templates. We analyze the properties \texttt{isExported},
\texttt{requiredModule} and \texttt{isConstructor} from the run-time JSON to distinguish
between \texttt{module-class} and \texttt{module-function}. If a function is invoked and
it is imported from the module we are analyzing we infer the template
\texttt{module-class} or \texttt{module-function}. We choose \texttt{module-class} if the
function is used as a constructor. Otherwise, we choose
\texttt{module-function}.
If no function is invoked, we use the \texttt{module}
template.
% \figref{fig:dts-generate-choose-templates} exposes this logic using the example
% provided in \figref{fig:multi-purpose-module-example}.

@@

We implemented only basic TypeScript features and we focused our effort in building an
end-to-end solution that generates valid and useful declaration files. Only the basic
types \texttt{string}, \texttt{number}, \texttt{boolean}, its union and the optional type were considered
both for function parameters and interface properties. Common types such as
\texttt{any}, arrays, function callbacks, tuples, intersections and features such as
generics, index signatures were not implemented.

Each declaration file in the DT repository was tagged with
\texttt{dts-parse}. We summarized the results in Table~\ref{tab:dts-parse-stats}. We based our decisions of prioritizing particular features on these results.

\begin{table}[tp]
  \begin{center}
    \begin{tabular}{ |c|c|c| } 
      \hline
      \textbf{TypeScript feature} & \textbf{Count Declaration Files} & \textbf{\%} \\ 
      \hline
      string & 5086 & 76.50\% \\
      optional-parameter & 4915 & 73.93\% \\
      boolean & 3891 & 58.53\% \\
      number & 3699 & 55.64\% \\
      void & 3548 & 53.37\% \\
      type-union & 3456 & 51.99\% \\
      type-function & 3286 & 49.43\% \\
      type-array & 3264 & 49.10\% \\
      type-any & 3127 & 47.04\% \\
      literals & 1925 & 28.96\% \\
      alias-type & 1899 & 28.56\% \\
      index-signature & 1271 & 19.12\% \\
      generics-function & 1145 & 17.22\% \\
      dot-dot-dot-token & 882 & 13.27\% \\
      call-signature & 817 & 12.29\% \\
      generics-interface & 718 & 10.80\% \\
      object & 661 & 9.94\% \\
      undefined & 577 & 8.68\% \\
      type-intersection & 431 & 6.48\% \\
      readonly & 373 & 5.61\% \\
      type-tuple & 355 & 5.34\% \\
      generics-class & 265 & 3.99\% \\
      static & 251 & 3.78\% \\
      private & 82 & 1.23\% \\
      public & 43 & 0.65\% \\
      protected & 19 & 0.29\% \\
      \hline
    \end{tabular}
  \end{center}
  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped. The 2nd column shows how many files in DT make use of a specific TypeScript feature. The 3rd column shows the corresponding percentage for a total of 6648 declaration files in DT. Numbers were obtained from commit \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
  \label{tab:dts-parse-stats}
\end{table}

>>>>

\title{Generation of TypeScript Declaration Files from JavaScript Code}
\subtitle{Catching the Developer's Intent from Examples}
% \titlerunning{Dummy short title}%optional, please use if title is longer than one line

@@

\author{Peter Thiemann}
\authorinfo{\email{thiemann@informatik.uni-freiburg.de}}
\affiliation{Albert-Ludwigs-Universität Freiburg, Germany}

@@

%TODO mandatory: add short abstract of the document
\begin{abstract}
Developers are starting to write large and complex applications in
TypeScript, a typed dialect of JavaScript. TypeScript applications
integrate JavaScript libraries via typed descriptions of their APIs
called declaration files. DefinitelyTyped is the standard public
repository for these files.
The repository is populated and maintained manually by volunteers, which
is error prone and time consuming. Discrepancies between a
declaration file and the JavaScript implementation lead to
incorrect feedback from the TypeScript IDE and thus to incorrect uses
of the underlying JavaScript library.

This work presents \texttt{dts-generate}, a tool that generates
TypeScript declaration files for JavaScript libraries uploaded to the \NPM{}
registry. It extracts code examples from the documentation written by
the developer, executes the library driven by the examples, gathers
run-time information, and generates a declaration file based on this
information. To evaluate the tool, 244 declaration files were
generated directly from an \NPM{} module and 82 of these were
compared with the corresponding declaration file provided on
DefinitelyTyped. We obtained positive results for 52 files, which
exhibited no differences at all or differences that can be resolved by
modifying the developer-provided examples. 
\end{abstract}

@@

\begin{lstlisting}[
    caption={\textbf{Unintuitive JavaScript behavior:} Falsy values,
    \lstinline!null! vs. \lstinline!undefined!,
 \lstinline!typeof!, and type coercion},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float=bp,
    captionpos=b
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3' 
typeof null; // object
null instanceof Object; // false
[1] + 1; //'11'
[2] == "2"; // true
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
\end{lstlisting}
% "01" < "00100"; // false 
% pjt: why unexpected?

@@

Unfortunately, creation and maintenance of most declaration
files in DefinitelyTyped is conducted manually,
which is time consuming and  error prone. Errors are aggravated because TypeScript takes a
declaration file at face value. Discrepancies between declaration and
implementing JavaScript library are not detected at 
compile time and result in incorrect behavior (e.g., autocompletion
hints) of the IDE. 
As TypeScript does not perform any run-time 
checking of types, these  discrepancies can lead to unexpected 
behavior and crashes. Such an experience can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

@@

We aim to improve on this situation by providing a tool that generates
sound TypeScript declaration files automatically. To do so, we rely on the examples provided
by the developer as part of the documentation of a library. If these
examples are sufficiently rich, then our toolchain can generate
high-quality declaration files for the library. This approach takes
advantage of 
best practices in the dynamic languages community which favors
examples and tests over writing out type signatures.

As an example, consider the NPM\footnote{NPM is the \href{https://www.npmjs.com/}{Node package
  manager}, an online repository for JavaScript modules.} module
\texttt{abs} that
``computes the absolute path of an input''. Its documentation consists
of the following three examples\footnote{\url{https://www.npmjs.com/package/abs}}.
\begin{lstlisting}[language=JavaScript,numbers=none]
const abs = require("abs");

@@

\item A comparator for TypeScript declaration files (Section~\ref{sec:dts-compare}). This tool is
  necessary for evaluating our framework and also useful to detect
  incompatibilities when evolving JavaScript modules.
\item An evaluation of our framework (Sections~\ref{sec:dts-generate-evaluation}
  and~\ref{sec:results}). We examined all 6029 entries in 
  the DefinitelyTyped repository and found 244 sufficiently
  well documented \NPM{} packages, on which we ran \texttt{dts-generate}
  and compared the outcome with the respective declaration file from the
  DefinitelyTyped repository. 
\end{itemize}

\section{Motivating Example}
\label{sec:motivating-example}
The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a
glob expression for matching filenames in the shell into a regular
expression\footnote{\url{https://www.npmjs.com/package/glob-to-regexp}}. It 
has about 9,900,000 weekly downloads and 188 \NPM{} packages depend on it. If
a developer creates or extends TypeScript code that depends on the
\texttt{glob-to-regexp} library, the TypeScript
compiler and IDE requires a declaration file for that library to
perform static checking and code completion, respectively. With
\texttt{dts-generate} we automatically generate a TypeScript
declaration file for \texttt{glob-to-regexp}. The tool downloads the \NPM{} 
package, runs the examples extracted from its documentation, gathers
run-time information, and generates a TypeScript declaration
file. The result is 
shown in \figref{fig:motivating-example-glob-to-regexp-vscode} and it is
ready for use in a TypeScript project. For example,  Visual Studio's
code completion runs properly with the generated file
(\figref{fig:motivating-example-glob-to-regexp-vscode}). If the
\texttt{glob-to-regexp} package gets modified in the future, a new declaration
file can be generated automatically using
\texttt{dts-generate}. Our comparator tool (Section~\ref{sec:dts-compare}) can compare the new file
for incompatibilities with the previous declaration file.

@@

The declaration file shown in \figref{fig:motivating-example-glob-to-regexp-vscode}
describes a package with a single exported function. The first parameter is of type
\texttt{string} and the second one is an optional object described by an interface. TypeScript namespaces
organize types declared in a declaration file and avoid name clashes with other types
\cite{typescript-namespaces}. For example, the interface
\lstinline[language=TypeScript]{I__opts} is declared in namespace
\lstinline[language=TypeScript]{globToRegexp} in
\figref{fig:motivating-example-glob-to-regexp-vscode}. Hence its uses
must be qualified by the namespace as in
\lstinline[language=TypeScript]{globToRegexp.I__opts}.

This file is an instance of one of the standard templates for writing
declaration
files\footnote{\url{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}}:
\textbf{module}, \textbf{module-class}, and
\textbf{module-function}. 
Each template corresponds to a different way of organizing the exports
of a JavaScript library. The choice of the template depends on the
structure of the underlying JavaScript library:
\begin{description}
\item[module] several exported functions,
\item[module-class] a class-like structure,
\item[module-function] exactly one exported function.
\end{description}
There are further templates, but most libraries fall in one of these
three categories.
Both libraries \texttt{glob-to-regexp} and \texttt{abs} are instances of the
\textbf{module-function} template.

\begin{lstlisting}[
    caption={Example for \textbf{module-function} template},
    language=bash,
	label=code:dts-generate-example,
    float,captionpos=b
]
$ ./dts-generate abs
$ cat output/abs/index.d.ts 
export = Abs;

@@

greet({
  greeting: "hello world",
  color: "#00ff00"
});
    \end{lstlisting}
    \caption{Example for \textbf{module-function} template}
    \label{fig:example-module-function}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=TypeScript,numbers=none]
export = GreetSettingsModule;

@@

}
    \end{lstlisting}
    \caption{Declaration file for \textbf{module-function} template}
    \label{fig:template-module-function}
  \end{subfigure}

@@

var goodbye = myLib.makeGoodBye();
console.log("The computed goodbye is: " + goodbye);    
      \end{lstlisting}
    \caption{Example for \textbf{module} template}
    \label{fig:example-module}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript,numbers=none]
export function makeGreeting(str: string): string;
export function makeGoodBye(): string;        
      \end{lstlisting}
      \caption{Declaration file for \textbf{module} template}
      \label{fig:template-module}
    \end{subfigure}

@@

var myGreeter = new Greeter("hello, world");
myGreeter.greeting = "howdy";
myGreeter.showGreeting();
      \end{lstlisting}
      \caption{Example for \textbf{module-class} template}
      \label{fig:example-class}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript,numbers=none]
export = Greeter;

@@

declare namespace Greeter {
}
      \end{lstlisting}
      \caption{Declaration file for \textbf{module-class} template}
      \label{fig:template-class}
    \end{subfigure}

@@

Like \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
file, we determine the kind of declaration file by examining the usage
of the module. If the
imported entity is used solely as a function as in Figure~\ref{fig:example-module-function}, the
generated declaration file follows the \textbf{module-function} template as shown in Figure~\ref{fig:template-module-function}.
If the example only accesses properties of the imported entity as in
Figure~\ref{fig:example-module}, we generate according to the \textbf{module} template (Figure~\ref{fig:template-module}).
If the imported entity is used with \lstinline/new/ to create new instances as in
Figure~\ref{fig:example-class}, we generate a \textbf{module-class} template (Figure~\ref{fig:template-class}).
While it is possible to create a library of undetermined category, our selection is driven
by the examples in the documentation and 
hence reflects the developers intent.
% Run-time analysis proves useful in this case,
% since it is not sufficient to statically analyze the exported module.

@@

The examples and the code base of the library are instrumented with
Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
information and type information at run time. Jalangi is a configurable
framework for dynamic analysis of JavaScript. It provides several
analysis modules that we extended as needed to retrieve the required
run-time information.
%, which is then saved in a JSON file.

A second independent block uses the run-time information to generate a
TypeScript declaration file. It infers the overall structure of the JavaScript
library, its interfaces, and the types from the run-time information. 
The resulting declaration file is ready for use in the development
process. Its contents mimic the usage of the library in the example
code fragments and match the structure of the
JavaScript library under analysis, so that the JavaScript code
generated after compiling the TypeScript code runs without
interface-related errors.

The command-line interface is very simple and inspired by the \texttt{dts-gen} tool
\cite{dts-gen} (see \coderef{code:dts-generate-example}). The only required
argument is the name of the module published to the \NPM{} registry.

@@

To extract run-time information from a JavaScript library, it is
necessary, by definition, to execute the code, because the
analysis modules provided by Jalangi can only gather information if the instrumented code gets executed.

@@

Option~\ref{item:1} does not solve our problem, it just delegates it
to the importing library, which also needs code to drive it. Moreover,
we need to identify a dependent package, which is costly to download and instrument.

We considered option~\ref{item:2} under the assumption that most
libraries would come with test cases. However, there is no standard
for testing JavaScript code so that test cases were difficult to reap
from the \NPM{} packages: they use different directory structures,
employ differing (or no) testing tools, or do not have tests
at all. Moreover, developers try to cover edge cases or to trigger errors by using illegal
input values in their tests. Such tests are not really useful for describing a library
from the consumer's point of view. For example, a developer might write a test invoking a
function with \texttt{null} just to validate that it throws an error
in such a scenario.

In the end, option~\ref{item:3} was the most viable option even though there
is no standard for documentation, either. However, almost all repositories
contain README files where the library authors briefly describe in prose what
the code does, which problem it solves, how to install the
application, how to build the code, etc. It is very common that
developers provide code examples in the README files to show how the
library works and how to use it. These code fragments showcase common use cases rather
than stress-testing the implementation (the problem of
option~\ref{item:2}) because they are meant to be instructive to users of the package.

@@

\subsection{Run-time Information Gathering}
\label{sec:run-time-information}
The Runtime Information block in
\figref{fig:tsd_generation_method_block_diagram} gathers
type-related information as well as usage-related information for each function parameter.
For example,

@@

Instead of analyzing the shape of the exported module, we choose the
template based on how the module is used. We analyze the properties \texttt{isExported},
\texttt{requiredModule}, and \texttt{isConstructor} from the
information gathered at run time to distinguish
between \textbf{module-class} and \textbf{module-function}. If a function is invoked and
it is imported from the module we are analyzing we infer the template
\textbf{module-class} or \textbf{module-function}. We choose \textbf{module-class} if the
function is used as a constructor. Otherwise, we choose
\textbf{module-function}.
If no function is invoked, we use the \textbf{module}
template.
% \figref{fig:dts-generate-choose-templates} exposes this logic using the example
% provided in \figref{fig:multi-purpose-module-example}.

@@

We implemented only basic TypeScript features and we focused our effort in building an
end-to-end solution that generates valid and useful declaration files. Only the basic
types \texttt{string}, \texttt{number}, \texttt{boolean}, unions
thereof, and optional elements were considered
both for function parameters and interface properties. We also
considered overloaded functions.
Common types such as
\texttt{any}, arrays, function callbacks, tuples, intersections (that
is, types formed using the intersection operator \lstinline[language=TypeScript]/&/) and features such as
generics, index signatures were not implemented.

Our selection of type features covered in the tool is driven by
examination of the files in
the DefinitelyTyped repository. We parsed each declaration file with
\texttt{dts-parse} and tagged it according to the features used in the file.
Table~\ref{tab:dts-parse-stats} summarizes the results.
We decided to prioritize the features with the highest number of
occurrences and support them first in our framework.

\begin{table}[tp]
  \begin{center}
    \begin{tabular}{ |l|c|r| } 
      \hline
      \textbf{TypeScript feature} & \textbf{Count} & \multicolumn{1}{|c|}{\textbf{\%}} \\ 
      \hline
      type-string & 5086 & 76.50\% \\
      optional-parameter & 4915 & 73.93\% \\
      type-boolean & 3891 & 58.53\% \\
      type-number & 3699 & 55.64\% \\
      type-void & 3548 & 53.37\% \\
      type-union & 3456 & 51.99\% \\
      type-function & 3286 & 49.43\% \\
      type-array & 3264 & 49.10\% \\
      type-any & 3127 & 47.04\% \\
      type-literals & 1925 & 28.96\% \\
      alias-type & 1899 & 28.56\% \\
      index-signature & 1271 & 19.12\% \\
      generics-function & 1145 & 17.22\% \\
      dot-dot-dot-token & 882 & 13.27\% \\
      call-signature & 817 & 12.29\% \\
      generics-interface & 718 & 10.80\% \\
      type-object & 661 & 9.94\% \\
      type-undefined & 577 & 8.68\% \\
      type-intersection & 431 & 6.48\% \\
      readonly & 373 & 5.61\% \\
      type-tuple & 355 & 5.34\% \\
      generics-class & 265 & 3.99\% \\
      static & 251 & 3.78\% \\
      private & 82 & 1.23\% \\
      public & 43 & 0.65\% \\
      protected & 19 & 0.29\% \\
      \hline
    \end{tabular}
  \end{center}
  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped (DT). The 2nd column shows how many files in DT make use of a specific TypeScript feature. The 3rd column shows the corresponding percentage for a total of 6648 declaration files in DT. Numbers were obtained from commit \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
  \label{tab:dts-parse-stats}
\end{table}
